---
title: "Highlander: From the Valleys to the Peak"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: |
  %\VignetteIndexEntry{Mock Fitting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Making Mock Data

Load the libraries we will need:

```{r}
library(ProSpect)
library(Highlander)
library(foreach)
library(celestial)
library(magicaxis)
```

Load some data we want to use:

```{r}
data("BC03lr") #BC03 spectral library
data("Dale_NormTot") #Normalised Dale templates
data("pivwave") # Pivot/effective wavelengths
```

```{r}
set.seed(1)
redshift = 0.1

filters=c('FUV_GALEX', 'NUV_GALEX', 'u_SDSS', 'g_SDSS', 'r_SDSS', 'i_SDSS', 'Z_VISTA',
          'Y_VISTA', 'J_VISTA', 'H_VISTA', 'K_VISTA', 'W1_WISE' , 'W2_WISE', 'W3_WISE',
          'W4_WISE', 'P100_Herschel', 'P160_Herschel', 'S250_Herschel' , 'S350_Herschel',
          'S500_Herschel')

filtout=foreach(i = filters)%do%{approxfun(getfilt(i))}

temppiv=pivwave[pivwave$filter %in% filters,]

agemax = 13.3e9-cosdistTravelTime(z=redshift, H0 = 67.8, OmegaM = 0.308)*1e9
```

Create our model galaxy We will be using a snorm_trunc SFH function and a massmap_lin metallicity with the following parameters:

```{r}
inpar=c(mSFR = 0, #log-space
        mpeak = 0.7, #log-space
        mperiod = 0.3, #log-space
        mskew = 0.3,
        tau_birth = 0, #log-space
        tau_screen = -0.5, #log-space
        alpha_SF_birth = 1,
        alpha_SF_screen = 3
        )
```

Simple function to view our target SFH:

```{r}
plotSFH=function(par,agemax=13.3, add=FALSE,col='black',ylim=NULL,...){
 magcurve(massfunc_snorm_trunc(age=x,mSFR=10^par[1],mpeak=10^par[2],mperiod=10^par[3],
                               mskew=par[4], magemax=agemax),0,13.8e9,add=add,col=col,
                               ylim=ylim,xlab='Age (Yr)', ylab='SFR (Msol / Yr)',...)
}
```

Let's take a look:

```{r, fig.width=8, fig.height=6}
plotSFH(inpar)
```

Now we can generate our galaxies SED:

```{r}
genSED=ProSpectSED(massfunc=massfunc_snorm_trunc,
                   mSFR=10^inpar[1],
                   mpeak=10^inpar[2],
                   mperiod=10^inpar[3],
                   mskew=inpar[4],
                   tau_birth=10^inpar[5], 
                   tau_screen=10^inpar[6], 
                   alpha_SF_birth=inpar[7], 
                   alpha_SF_screen=inpar[8],
                   z=0.1,
                   Z=Zfunc_massmap_lin,
                   filtout=filtout,
                   Dale=Dale_NormTot,
                   speclib=BC03lr,
                   agemax=agemax
)
```

At this point we create our mock photometry catalogue entry with a fractional error in flux of 0.1 assumed (roughly 0.092 mag error). The photom output of **ProSpectSED** is always Jansky, which is what we want for fitting (you shouldn't fit in magnitude space!).

```{r}
flux_input=data.frame(filter=temppiv$filter, pivwave=temppiv$pivwave, flux=genSED$Photom, fluxerr=genSED$Photom*0.1)
print(flux_input)
```

## Fitting a Galaxy

In principle if you have a real galaxy to fit and it is already in the required format of **flux_input**, you can start from this point.

To speed things up we will also pre-compute the Luminosity distance to our source (since this will be the same for every iteration, and will save some computation time):

```{r}
LumDist_Mpc = cosdistLumDist(z=0.1, H0 = 67.8, OmegaM = 0.308)
```

```{r}
Data = list(flux=flux_input,
          arglist=list(z=0.1, massfunc=massfunc_snorm_trunc, agemax=agemax, Z=Zfunc_massmap_lin, LumDist_Mpc=LumDist_Mpc),
          speclib=BC03lr, 
          Dale=Dale_NormTot, 
          filtout=filtout, 
          SFH=SFHfunc, # the preferred functional form of the SFH (eg either SFHfunc, SFHburst)
          parm.names=c('mSFR','mpeak','mperiod','mskew','tau_birth','tau_screen',
                       'alpha_SF_birth','alpha_SF_screen'), # which parameters to fit for
          logged=c(T,T,T,F,T,T,F,F), # fit parameters in logged or linear space
          fit = 'cma', # specifies the way in which the SED should be fitted ('LD', 'optim', 'CMA', or 'check')
          N=length(filters), # number of observed filters
          like='norm',
          verbose=FALSE
  )
```

Now we are going to use **Lowlander** to find reasonble paramter limits for our fitting:

```{r}
library(lhs)

lower = c(-4,-2,-1,-0.5,-2.5,-2.5,0,0)
upper = c(3,1,1,1,1.5,1,4,4) # initial range for parameters

# Note for Lowlander the only output from our likelihood function has to be LL of ChiSq:
intervals = Lowlander(lower, upper, Data=Data, likefunc=ProSpectSEDlike)

intervals$vol_red #the volume reduction factor, which is how much less hyper-space we wille explore.
lower = intervals$lower
upper = intervals$upper
Data$fit = 'LD' # now we can switch it to full LD mode.
```

Now we can run an MCMC fit using **Highlander** (depending on machine, this might take 1-2 minutes):

```{r eval=FALSE}
badpar = (lower + upper) / 2
High_out = Highlander(badpar, Data, ProSpectSEDlike, lower=lower, upper=upper,
                      Niters=c(200,200))
```

```{r eval=FALSE}
Data$fit = 'check'
check_out = ProSpectSEDlike(High_out$parm, Data)
plot(check_out)
plot(check_out$SEDout)
plotSFH(inpar)
lines(check_out$SEDout$Stars$agevec, check_out$SEDout$Stars$SFR, col='red')
```

## Multi-Core Fitting

**Highlander** supports parallel optimisation via the `cores` argument. Setting `cores > 1` launches that many independent **Highlander** chains simultaneously, each starting from a different random seed (`seed`, `seed+1`, ..., `seed+cores-1`). All chains explore the likelihood surface independently, and the chain that achieves the highest log-posterior (LP) is returned. This is particularly useful for multi-modal problems where a single chain may get trapped in a local optimum.

On Unix-like systems (Linux, macOS) `parallel::mclapply` is used, which forks the current R process with no extra setup. On Windows a PSOCK cluster is used automatically — no changes to your code are needed when moving between platforms.

```{r eval=FALSE}
# Run 4 independent chains in parallel and keep the best result.
# On a 4-core machine this takes roughly the same wall time as a single chain.
High_out_mc = Highlander(badpar, Data, ProSpectSEDlike, lower=lower,
                         upper=upper, Niters=c(200,200), cores=4L)

# The returned object is identical in structure to the single-core output.
# $parm holds the best parameters found across all parallel chains.
High_out_mc$parm
High_out_mc$LP

# best_job records which chain (1-indexed) achieved the highest LP.
High_out_mc$best_job
```

Use `keepall = TRUE` to retain the full output of every parallel chain. This is useful when you want to compare posterior distributions across chains, or diagnose convergence problems:

```{r eval=FALSE}
# Run 4 chains in parallel, keeping all outputs.
High_out_mc_all = Highlander(badpar, Data, ProSpectSEDlike, lower=lower, upper=upper,
                             Niters=c(200,200), cores=4L, keepall=TRUE)

# High_jobs is a list of length cores; each element is a standard Highlander output.
length(High_out_mc_all$High_jobs)  # 4

# best_job tells you which one won:
High_out_mc_all$best_job  # e.g. 2

# Directly access the winning chain (equivalent to High_out_mc_all itself, but useful for comparison):
best_chain = High_out_mc_all$High_jobs[[High_out_mc_all$best_job]]

# Or inspect a non-winning chain:
High_out_mc_all$High_jobs[[1]]$LP
```

```{r eval=FALSE}
# Inspect the best-fit SED from the parallel run, exactly as before.
Data$fit = 'check'
check_out_mc = ProSpectSEDlike(High_out_mc$parm, Data)
plot(check_out_mc)
plot(check_out_mc$SEDout)
plotSFH(inpar)
lines(check_out_mc$SEDout$Stars$agevec, check_out_mc$SEDout$Stars$SFR, col='red')
```
