---
title: "MCMC Algorithms in LaplacesDemon"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: |
  %\VignetteIndexEntry{MCMC Algorithms in LaplacesDemon}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Introduction

Markov Chain Monte Carlo (MCMC) is the engine that drives Bayesian
posterior inference in **LaplacesDemon**.  The main entry point —
`LaplacesDemon()` — exposes a rich collection of MCMC algorithms through
its `Algorithm` argument.  Choosing the right algorithm for a given
problem can dramatically affect mixing, convergence speed, and
computational cost.

This vignette provides a comprehensive, mathematically-oriented
description of every algorithm currently available via the `Algorithm`
argument, together with practical guidance on when and how to use each
one.  The algorithms are grouped by conceptual family to make it easier
to navigate the large menu.

Much of the material and explanations for the algorithms was taken from
the snapshot of the old Statisticat (creators of **LaplacesDemon**) website on
Bayesian Inference captured on the Way Back Machine (see [Bayesian Inference](https://web.archive.org/web/20150128091453/http://www.bayesian-inference.com/index)).

## Notation

Throughout this document we use the following notation:

- $\theta \in \mathbb{R}^d$ — the parameter vector of dimension $d$.
- $\pi(\theta)$ — the unnormalised target (posterior) density, evaluated
  on the log scale as $\log\pi(\theta)$ (`LP` in **LaplacesDemon**).
- $q(\theta' \mid \theta)$ — the proposal density.
- $\alpha(\theta, \theta')$ — the Metropolis–Hastings acceptance
  probability.
- $U(\theta) = -\log\pi(\theta)$ — the potential energy function (used
  in Hamiltonian samplers).
- $K(p) = \frac{1}{2}p^T M^{-1} p$ — the kinetic energy, where $p$ is
  the auxiliary momentum and $M$ is a mass matrix.

## Using the `Algorithm` Argument

The algorithm is selected by passing a character string to the
`Algorithm` argument of `LaplacesDemon()`:

```{r demo-usage}
library(LaplacesDemon)

Fit <- LaplacesDemon(
  Model,
  Data        = MyData,
  Initial.Values = iv,
  Iterations  = 10000,
  Thinning    = 10,
  Algorithm   = "CHARM",               # <-- select algorithm here
  Specs       = list(alpha.star = NA)  # <-- algorithm-specific specs
)
```

Each algorithm may require a different `Specs` list.  The required and
optional elements of `Specs` for every algorithm are documented below.

---

# Metropolis–Hastings Algorithms

The Metropolis–Hastings (MH) framework (Metropolis et al. 1953; Hastings 1970)
forms the foundation for the largest family of algorithms in
**LaplacesDemon**.  A proposal $\theta'$ is drawn from
$q(\theta' \mid \theta^{(t)})$ and accepted with probability

$$
\alpha = \min\!\left(1,\;
  \frac{\pi(\theta')\,q(\theta^{(t)} \mid \theta')}
       {\pi(\theta^{(t)})\,q(\theta' \mid \theta^{(t)})}
\right).
$$

When $q$ is symmetric the ratio $q(\theta^{(t)} \mid \theta') /
q(\theta' \mid \theta^{(t)})$ drops out, giving the original Metropolis
acceptance ratio.

## Random-Walk Metropolis (`RWM`)

**Description.**  The simplest MH variant.  A zero-mean, isotropic
Gaussian perturbation is added to the current state:

$$
\theta' = \theta^{(t)} + \epsilon, \qquad
\epsilon \sim \mathcal{N}(0,\,\Sigma).
$$

For well-tuned proposals the optimal acceptance rate is approximately
23.4% for high-dimensional targets (Roberts et al. 1997).

**Specs.**

| Spec | Description |
|------|-------------|
| `B`  | Optional list of parameter blocks. Each block is updated as a group. |

**When to use.**  Good default choice for low- to moderate-dimensional
problems when no gradient information is available.  Requires careful
tuning of $\Sigma$ for efficiency.

```{r rwm}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "RWM",
  Specs     = list(B = NULL))
```

## Independence Metropolis (`IM`)

**Description.**  Proposals are drawn from a fixed distribution
$q(\theta')$ that does not depend on the current state.  This is
equivalent to importance sampling within an MH accept/reject step.  It
is most efficient when $q$ closely approximates $\pi$, and can be used
with a multivariate normal or t proposal centred on the posterior mode.

**Specs.**

| Spec | Description |
|------|-------------|
| `mu` | Mean of the multivariate proposal distribution (usually the posterior mode). |

**When to use.**  Useful when a good approximation to $\pi$ is
available (e.g., from `LaplaceApproximation()`).  Poor mixing when $q$
deviates substantially from $\pi$.

```{r im}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "IM",
  Specs     = list(mu = posterior_mode))
```

## Metropolis-within-Gibbs (`MWG`)

**Description.**  Each parameter (or block) is updated in turn with a
univariate Metropolis–Hastings step, conditional on all other
parameters.  The proposal variance for each component is tuned
adaptively to reach a target acceptance rate of approximately 44%.
This is the default algorithm in **LaplacesDemon**.

**Specs.**

| Spec | Description |
|------|-------------|
| `B`  | Optional list of blocks.  If `NULL`, each parameter is its own block. |

**When to use.**  Very robust; works well even with correlated
parameters because it updates one dimension at a time.  May be slow for
strongly correlated, high-dimensional posteriors because it cannot
exploit inter-parameter correlations within a single proposal.

```{r mwg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "MWG",
  Specs     = list(B = NULL))
```

## Adaptive Metropolis-within-Gibbs (`AMWG`)

**Description.**  Extends MWG with componentwise adaptation of the
log-scale proposal standard deviations following Roberts and Rosenthal
(2009).  Every `Periodicity` iterations the log-scale step is
incremented or decremented to push each component's acceptance rate
towards 44%.

**Specs.**

| Spec | Description |
|------|-------------|
| `B`           | Optional blocks. |
| `n`           | Number of previous iterations (for continuing a run). |
| `Periodicity` | How often to adapt (iterations). |

```{r amwg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AMWG",
  Specs     = list(B = NULL, n = 0, Periodicity = 50))
```

## Adaptive Metropolis (`AM`)

**Description.**  The Adaptive Metropolis algorithm of Haario, Saksman
and Tamminen (2001) learns a full multivariate Gaussian proposal
covariance from the history of the chain.  After a burn-in of
`Adaptive` iterations the proposal covariance is updated every
`Periodicity` iterations:

$$
\Sigma_t = s_d \left[\operatorname{Cov}(\theta^{(0)},\ldots,\theta^{(t-1)}) + \varepsilon I_d\right],
$$

where $s_d = 2.38^2/d$ is the theoretically optimal scaling for a
Gaussian target (Gelman et al. 1996).

**Specs.**

| Spec | Description |
|------|-------------|
| `Adaptive`    | Iteration at which adaptation begins. |
| `Periodicity` | Adaptation frequency in iterations. |

```{r am}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AM",
  Specs     = list(Adaptive = 500, Periodicity = 10))
```

## Adaptive-Mixture Metropolis (`AMM`)

**Description.**  A mixture of an adaptive multivariate proposal
(similar to AM) and a static componentwise proposal.  Each iteration,
the componentwise proposal is used with probability $w$ and the
adaptive multivariate proposal with probability $1-w$.  This helps
maintain ergodicity while benefiting from adaptation
(Roberts & Rosenthal 2009).

**Specs.**

| Spec | Description |
|------|-------------|
| `Adaptive`    | Iteration at which adaptation begins. |
| `B`           | Optional blocks for the componentwise component. |
| `n`           | Previous iterations for continuing a run. |
| `Periodicity` | Adaptation frequency. |
| `w`           | Mixture weight for componentwise proposals (recommended: 0.05). |

```{r amm}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AMM",
  Specs     = list(Adaptive = 500, B = NULL, n = 0,
                   Periodicity = 10, w = 0.05))
```

## Delayed Rejection Metropolis (`DRM`)

**Description.**  When a proposal is rejected, a second-stage proposal
of smaller scale is attempted before the chain remains at the current
state.  This reduces the probability of long sequences of rejections
without violating detailed balance (Mira 2001).

**Specs.**  `DRM` requires no `Specs` (pass `NULL` or omit).

```{r drm}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "DRM",
  Specs     = NULL)
```

## Delayed Rejection Adaptive Metropolis (`DRAM`)

**Description.**  Combines delayed rejection (DRM) with the adaptive
multivariate proposal of AM.  After `Adaptive` iterations the proposal
covariance is updated every `Periodicity` iterations, and rejected
proposals trigger a second, smaller attempt (Haario et al. 2006).

**Specs.**

| Spec | Description |
|------|-------------|
| `Adaptive`    | Start of adaptation. |
| `Periodicity` | Adaptation frequency. |

```{r dram}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "DRAM",
  Specs     = list(Adaptive = 500, Periodicity = 10))
```

## Robust Adaptive Metropolis (`RAM`)

**Description.**  Vihola's (2011) RAM algorithm adapts the proposal
Cholesky factor $S$ so that the mean acceptance rate converges to a
target $\alpha^*$.  The adaptation rule is:

$$
S_{t+1} S_{t+1}^T = S_t \!\left(I + \eta_t
  (\alpha_t - \alpha^*) \frac{u_t u_t^T}{\|u_t\|^2}\right) S_t^T,
$$

where $u_t = \theta' - \theta^{(t)}$ and $\eta_t = \min(1, d\,t^{-\gamma})$.
RAM is guaranteed to converge under mild conditions.

**Specs.**

| Spec | Description |
|------|-------------|
| `alpha.star` | Target acceptance rate (recommended 0.234 for multivariate). |
| `B`          | Optional blocks. |
| `Dist`       | Proposal distribution: `"N"` (normal) or `"t"`. |
| `gamma`      | Decay rate of adaptation, in $(0.5, 1]$ (recommended 0.66). |
| `n`          | Previous iterations. |

```{r ram}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "RAM",
  Specs     = list(alpha.star = 0.234, B = NULL,
                   Dist = "N", gamma = 0.66, n = 0))
```

## Multiple-Try Metropolis (`MTM`)

**Description.**  Instead of a single proposal, $K$ candidate points
are drawn from the proposal distribution.  One candidate $\theta'$ is
selected with probability proportional to $\pi(\theta')\,w(\theta',\theta)$,
and accepted via a modified MH ratio that maintains detailed balance
(Liu et al. 2000).  For a fixed computational budget it can achieve better
mixing than standard MH.

**Specs.**

| Spec | Description |
|------|-------------|
| `K`        | Number of proposals per iteration (integer ≥ 1). |
| `CPUs`     | Number of CPUs for parallel proposal evaluation (≥ 1). |
| `Packages` | Packages to load on each CPU (`NULL` if none). |
| `Dyn.libs` | Dynamic libraries to load on each CPU (`NULL` if none). |

```{r mtm}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "MTM",
  Specs     = list(K        = 4,
                   CPUs     = 1,
                   Packages = NULL,
                   Dyn.libs = NULL))
```

## Random Dive Metropolis-Hastings (`RDMH`)

**Description.**  A multiplicative random-walk sampler on $\mathbb{R}$,
proposed by Dutta (2012).  The proposal is
$\theta'_j = \theta_j \cdot \exp(\epsilon_j)$, $\epsilon_j \sim
\mathcal{N}(0, \sigma^2)$.  This makes it scale-free and avoids the
sign problem of additive proposals when parameters are constrained to be
positive.

**Specs.**  `RDMH` requires no `Specs` (pass `NULL` or omit).

```{r rdmh}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "RDMH",
  Specs     = NULL)
```

---

# Componentwise and Hit-and-Run Samplers

These algorithms move in one direction at a time but vary *which*
direction is chosen at each step.

## Componentwise Hit-And-Run Metropolis (`CHARM`)

**Description.**  The default algorithm used by **Highlander** when
calling `LaplacesDemon()`.  At each iteration a random direction $v$ is
drawn uniformly from the unit sphere and a step of size $\sigma$ is
proposed along $v$:

$$
\theta' = \theta^{(t)} + \sigma v.
$$

Unlike standard MWG, CHARM mixes coordinate directions, which can aid
mixing when parameters are correlated.  When `alpha.star` is provided,
the scale $\sigma$ is tuned to approach that acceptance rate.

**Specs.**

| Spec | Description |
|------|-------------|
| `alpha.star` | Optional target acceptance rate (recommended 0.44 for componentwise). Set to `NA` to disable adaptive scaling. |

```{r charm}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "CHARM",
  Specs     = list(alpha.star = NA))
```

## Hit-And-Run Metropolis (`HARM`)

**Description.**  Similar to CHARM but the direction is drawn from a
fixed (non-adapting) distribution.  Smith (1984) originally proposed
hit-and-run for uniform sampling over convex sets; the HARM variant
adapts it to general log-concave targets using Gaussian proposals along
the chosen direction.

**Specs.**

| Spec | Description |
|------|-------------|
| `alpha.star` | Optional target acceptance rate (recommended 0.234 for multivariate). Set to `NA` to disable adaptive scaling. |
| `B`          | Optional list of blocks. |

```{r harm}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "HARM",
  Specs     = list(alpha.star = NA, B = NULL))
```

## Adaptive Directional Metropolis-within-Gibbs (`ADMG`)

**Description.**  Bai (2009) proposed updating each parameter along an
adaptively-chosen direction within the Gibbs framework.  The direction
for each component is adapted using past samples to align with the local
curvature of the posterior, reducing the correlations that slow
componentwise MH.

**Specs.**

| Spec | Description |
|------|-------------|
| `n`           | Number of previous iterations (for continuing a run). |
| `Periodicity` | How often to adapt (iterations); defaults to `length(Initial.Values)`. |

```{r admg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "ADMG",
  Specs     = list(n = 0, Periodicity = 1))
```

---

# Gibbs Samplers

Gibbs sampling draws each parameter (or block) exactly from its full
conditional distribution $\pi(\theta_j \mid \theta_{-j}, y)$.  When
full conditionals are available in closed form, Gibbs is highly
efficient.

## Gibbs Sampler (`Gibbs`)

**Description.**  The user supplies a function `FC` that computes and
returns updated parameters from their full conditionals.  This is the
exact Gibbs sampler of Geman and Geman (1984); no Metropolis–Hastings
correction is needed because samples are drawn exactly.

**Specs.**

| Spec | Description |
|------|-------------|
| `FC`  | A function taking `(parm, Data)` and returning an updated parameter vector. |
| `MWG` | Integer vector of parameter indices to receive MWG updates instead of exact Gibbs updates. |

```{r gibbs}
FC <- function(parm, Data) {
  # draw each parameter from its full conditional
  parm[1] <- rnorm(1, ...)
  return(parm)
}

Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "Gibbs",
  Specs     = list(FC = FC, MWG = NULL))
```

## Griddy-Gibbs (`GG`)

**Description.**  When full conditionals are not available analytically,
Ritter and Tanner (1992) proposed evaluating the unnormalised
conditional on a user-defined grid and drawing via inverse CDF.  This
is exact up to the grid resolution and avoids Metropolis correction.

**Specs.**

| Spec | Description |
|------|-------------|
| `dparm`    | Vector of indices of discrete parameters (`NULL` for all continuous). |
| `Grid`     | A vector (or list of vectors) of evenly-spaced grid points. |
| `CPUs`     | Number of CPUs for parallel evaluation (≥ 1). |
| `Packages` | Character vector of packages to load on each CPU (`NULL` if none). |
| `Dyn.libs` | Character vector of dynamic libraries to load on each CPU (`NULL` if none). |

```{r gg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "GG",
  Specs     = list(dparm    = NULL,
                   Grid     = seq(-0.1, 0.1, length.out = 5),
                   CPUs     = 1,
                   Packages = NULL,
                   Dyn.libs = NULL))
```

## Adaptive Griddy-Gibbs (`AGG`)

**Description.**  Extends GG by adapting the grid width using the
running variance of each parameter.  This removes the burden of
hand-tuning the grid while preserving the exact-Gibbs character of the
sampler.

**Specs.**

| Spec | Description |
|------|-------------|
| `dparm`    | Discrete parameter indices (`NULL` for all continuous). |
| `Grid`     | Initial grid vector or list of vectors (adapted subsequently). |
| `smax`     | Maximum allowable grid width (standard deviation). |
| `CPUs`     | Number of CPUs for parallel evaluation (≥ 1). |
| `Packages` | Packages to load on each CPU (`NULL` if none). |
| `Dyn.libs` | Dynamic libraries to load on each CPU (`NULL` if none). |

```{r agg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AGG",
  Specs     = list(dparm    = NULL,
                   Grid     = seq(-0.1, 0.1, length.out = 5),
                   smax     = Inf,
                   CPUs     = 1,
                   Packages = NULL,
                   Dyn.libs = NULL))
```

---

# Ensemble and Population-Based Samplers

These methods maintain multiple "walkers" or "chains" simultaneously,
allowing inter-chain proposals that exploit the current distribution of
the ensemble.

## Affine-Invariant Ensemble Sampler (`AIES`)

**Description.**  Goodman and Weare's (2010) ensemble sampler maintains
$N$ walkers $\{\theta^{(k)}\}_{k=1}^N$.  For walker $k$, a complement
walker $j \ne k$ is selected at random and the proposal is

$$
\theta' = \theta^{(j)} + Z \left(\theta^{(k)} - \theta^{(j)}\right),
\quad Z \sim g(z),
$$

where $g$ is chosen so the sampler is affine-invariant.  The default
choice gives $Z$ with PDF $g(z) \propto z^{-1/2}$ on $[1/\beta, \beta]$
with $\beta = 2$.

**Specs.**

| Spec | Description |
|------|-------------|
| `Nc`       | Number of walkers (≥ 3; must be even when `CPUs > 1`). |
| `Z`        | Optional $N_c \times d$ matrix of initial walker positions (`NULL` to auto-generate). |
| `beta`     | Scale parameter controlling the stretch move (default 2; must be > 1). |
| `CPUs`     | Number of CPUs for parallel walker updates (≥ 1). |
| `Packages` | Packages to load on each CPU (`NULL` if none). |
| `Dyn.libs` | Dynamic libraries to load on each CPU (`NULL` if none). |

```{r aies}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AIES",
  Specs     = list(Nc       = 16,
                   Z        = NULL,
                   beta     = 2,
                   CPUs     = 1,
                   Packages = NULL,
                   Dyn.libs = NULL))
```

## Differential Evolution Markov Chain (`DEMC`)

**Description.**  DEMC (Ter Braak & Vrugt 2008) runs $N_c \ge 3$ chains in
parallel.  The proposal for chain $i$ uses the difference of two
randomly-chosen other chains:

$$
\theta'_i = \theta_i^{(t)} + \gamma (\theta_r^{(t)} - \theta_s^{(t)}) + e,
\quad e \sim \mathcal{N}(0, \epsilon I),
$$

where $\gamma = 2.38/\sqrt{2d}$ by default.  A "snooker" update
(probability $w$) projects along the line connecting the current chain
to a randomly-chosen history point.

**Specs.**

| Spec | Description |
|------|-------------|
| `gamma` | Scale for difference vector (defaults to $2.38/\sqrt{2d}$). |
| `Nc`    | Number of parallel chains (≥ 3). |
| `w`     | Probability of snooker update (recommended 0.1). |
| `Z`     | Optional matrix/array of previous thinned samples. |

```{r demc}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "DEMC",
  Specs     = list(gamma = NULL, Nc = 3, w = 0.1, Z = NULL))
```

## Interchain Adaptation (`INCA`)

**Description.**  Craiu, Rosenthal and Yang (2009) proposed INCA, which
uses parallel chains to adapt a shared proposal covariance.  Regional
information from nearby chains is combined to improve the proposal for
each chain, yielding faster mixing than independent AM chains.

**Specs.**

| Spec | Description |
|------|-------------|
| `Adaptive`    | Start of adaptation. |
| `Periodicity` | Adaptation frequency. |

```{r inca}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "INCA",
  Specs     = list(Adaptive = 500, Periodicity = 10))
```

## Metropolis-Coupled MCMC (`MCMCMC`)

**Description.**  Geyer's (1991) parallel tempering runs $T$ chains at
different temperatures $\{1/\lambda_t\}$.  The cold chain ($t=1$)
targets $\pi(\theta)$; hotter chains explore more freely.  Periodically,
adjacent chains attempt a swap:

$$
\alpha_{\text{swap}} = \min\!\left(1,\;
  \frac{\pi(\theta_t)^{1/\lambda_{t+1}} \pi(\theta_{t+1})^{1/\lambda_t}}
       {\pi(\theta_t)^{1/\lambda_t}   \pi(\theta_{t+1})^{1/\lambda_{t+1}}}
\right).
$$

This helps the cold chain escape local modes.

**Specs.**

| Spec | Description |
|------|-------------|
| `lambda` | Temperature spacing: scalar or vector of temperatures. |
| `CPUs`   | Number of CPUs (must be ≥ 2; one per temperature). |
| `Packages` | Packages to load on each CPU (`NULL` if none). |
| `Dyn.libs` | Dynamic libraries to load on each CPU (`NULL` if none). |

```{r mcmcmc}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "MCMCMC",
  Specs     = list(lambda   = 1,
                   CPUs     = 2,
                   Packages = NULL,
                   Dyn.libs = NULL))
```

---

# Hamiltonian / Langevin Algorithms

Hamiltonian Monte Carlo (HMC) (Duane et al. 1987; Neal 2011) augments $\theta$
with an auxiliary momentum $p \sim \mathcal{N}(0, M)$ and simulates
Hamiltonian dynamics

$$
\frac{d\theta}{dt} = \frac{\partial K}{\partial p},
\qquad
\frac{dp}{dt} = -\frac{\partial U}{\partial \theta},
$$

using the leapfrog integrator.  Proposals obtained by integrating for
$L$ steps of size $\epsilon$ are accepted/rejected by the energy
$H(\theta, p) = U(\theta) + K(p)$.  Because dynamics preserve volume
and (approximately) energy, acceptance rates are high and proposals are
far from the starting point.

## Hamiltonian Monte Carlo (`HMC`)

**Description.**  The classic HMC sampler.  Gradient
$\nabla_\theta \log\pi(\theta)$ must be evaluated at each leapfrog
step, making it expensive per iteration but typically much more
efficient per independent sample than gradient-free methods.

**Specs.**

| Spec | Description |
|------|-------------|
| `epsilon` | Leapfrog step size vector (length $d$). |
| `L`       | Number of leapfrog steps per iteration. |
| `m`       | Mass matrix ($d \times d$). |

```{r hmc}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "HMC",
  Specs     = list(epsilon = 0.1, L = 20, m = NULL))
```

## Adaptive Hamiltonian Monte Carlo (`AHMC`)

**Description.**  Adapts the mass matrix and step sizes from the running
covariance of the chain, removing the need to hand-tune these HMC
parameters.

**Specs.**

| Spec | Description |
|------|-------------|
| `epsilon`     | Initial step-size vector. |
| `L`           | Leapfrog steps. |
| `m`           | Initial mass matrix. |
| `Periodicity` | Adaptation frequency. |

```{r ahmc}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AHMC",
  Specs     = list(epsilon = 0.1, L = 20, m = NULL, Periodicity = 10))
```

## Hamiltonian Monte Carlo with Dual-Averaging (`HMCDA`)

**Description.**  Hoffman and Gelman's (2014) HMCDA uses Nesterov's
dual-averaging scheme to automatically tune the step size $\epsilon$ to
achieve a target acceptance rate $\delta$ (typically 0.65), and adapts
the trajectory length $\lambda = L\epsilon$ to a fixed user-supplied
target.

**Specs.**

| Spec | Description |
|------|-------------|
| `delta`    | Target acceptance rate (recommended 0.65). |
| `epsilon`  | Initial step size (scalar; or `NULL` for automatic initialisation). |
| `lambda`   | Target trajectory length. |
| `Lmax`     | Maximum number of leapfrog steps. |
| `A`        | Number of adaptive burn-in iterations to discard after dual-averaging. |

```{r hmcda}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "HMCDA",
  Specs     = list(A       = 500,
                   delta   = 0.65,
                   epsilon = NULL,
                   lambda  = 0.1,
                   Lmax    = 1000))
```

## No-U-Turn Sampler (`NUTS`)

**Description.**  NUTS (Hoffman & Gelman 2014) eliminates the need to choose $L$
by automatically stopping the leapfrog trajectory when it would start to
backtrack ("make a U-turn").  Combined with dual-averaging for $\epsilon$,
NUTS is nearly parameter-free and often outperforms hand-tuned HMC.

**Specs.**

| Spec | Description |
|------|-------------|
| `A`       | Number of adaptive burn-in iterations to discard. |
| `delta`   | Target acceptance rate (recommended 0.6). |
| `epsilon` | Initial step size (or `NULL`). |
| `Lmax`    | Maximum tree depth (maximum leapfrog steps = $2^{\text{Lmax}}$). |

```{r nuts}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "NUTS",
  Specs     = list(A = 500, delta = 0.6,
                   epsilon = NULL, Lmax = 1000))
```

## Tempered Hamiltonian Monte Carlo (`THMC`)

**Description.**  A variant of HMC that heats the momentum in the first
half of the leapfrog trajectory and cools it in the second half,
controlled by a `Temperature` parameter.  Higher temperatures encourage
exploration of multimodal posteriors.

**Specs.**

| Spec | Description |
|------|-------------|
| `epsilon`     | Leapfrog step-size vector. |
| `L`           | Leapfrog steps. |
| `m`           | Mass matrix. |
| `Temperature` | Tempering value (> 1 increases exploration). |

```{r thmc}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "THMC",
  Specs     = list(epsilon = 0.1, L = 20, m = NULL, Temperature = 1))
```

## Metropolis-Adjusted Langevin Algorithm (`MALA`)

**Description.**  MALA uses a single gradient-guided Langevin step

$$
\theta' = \theta^{(t)}
  + \frac{\epsilon^2}{2} \nabla\log\pi(\theta^{(t)}) + \epsilon\,\xi,
\quad \xi \sim \mathcal{N}(0, I),
$$

followed by an MH correction.  Atchadé (2006) proposed an adaptive
truncated version that clips the drift to prevent divergence.  The
optimal acceptance rate for MALA is approximately 57.4%.

**Specs.**

| Spec | Description |
|------|-------------|
| `A`          | Maximum acceptable norm for adaptive parameters. |
| `alpha.star` | Target acceptance rate (recommended 0.574). |
| `delta`      | Constant in the bounded drift (default 1; range $[10^{-10}, 1000]$). |
| `epsilon`    | Length-2 vector: (min adaptive scale, regularisation). |
| `gamma`      | In $(1, \text{Iterations})$; controls adaptation decay (default 1). |

```{r mala}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "MALA",
  Specs     = list(A = Inf, alpha.star = 0.574,
                   delta = 1, epsilon = c(1e-6, 1e-7),
                   gamma = 1))
```

## Stochastic Gradient Langevin Dynamics (`SGLD`)

**Description.**  Welling and Teh (2011) proposed SGLD for big-data
settings.  At each iteration, a random minibatch of `size` rows is
drawn from a large dataset stored on disk, and a noisy gradient is used
in a Langevin step with step size $\epsilon_t$.  No MH correction is
needed when $\epsilon_t \to 0$.

**Specs.**

| Spec | Description |
|------|-------------|
| `epsilon` | Step-size scalar or vector of length `Iterations`. |
| `file`    | Path to a `.csv` file of the full data. |
| `Nc`      | Number of columns in the data file. |
| `Nr`      | Number of rows in the data file. |
| `size`    | Minibatch size (rows per iteration). |

```{r sgld}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "SGLD",
  Specs     = list(epsilon = 1e-4, file = "bigdata.csv",
                   Nc = 10, Nr = 1e6, size = 200))
```

## Preconditioned Crank-Nicolson (`pCN`)

**Description.**  Beskos et al. (2008) introduced pCN for
infinite-dimensional targets.  The proposal is

$$
\theta' = \sqrt{1-\beta^2}\,\theta^{(t)} + \beta\,\xi,
\quad \xi \sim \mathcal{N}(0, C),
$$

where $C$ is a prior covariance.  pCN is dimension-robust: its
acceptance rate does not degrade as dimension increases, making it
suitable for large hierarchical models.

**Specs.**

| Spec | Description |
|------|-------------|
| `beta` | Autoregressive parameter, $\beta \in (0,1)$. |

```{r pcn}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "pCN",
  Specs     = list(beta = 0.1))
```

---

# Slice Samplers

Slice sampling (Neal 2003) avoids explicit proposals entirely by
sampling uniformly from the "slice" $\{(\theta, y): y \le \pi(\theta)\}$,
which is equivalent to sampling from $\pi$.  Various schemes differ in
how the slice interval is constructed.

## Slice Sampler (`Slice`)

**Description.**  Neal's (2003) general univariate (or blockwise) slice
sampler uses "stepping out" and "shrinkage" to find the slice.  For
each parameter (or block) an interval of width $w$ is grown until it
brackets the slice, then a point is drawn uniformly from the interval
and shrunk if necessary.

**Specs.**

| Spec | Description |
|------|-------------|
| `B`      | Optional blocks. |
| `Bounds` | Lower and upper bounds for the slice (vector or list). |
| `m`      | Maximum steps for growing the interval (integer ≥ 1). |
| `Type`   | `"Continuous"`, `"Nominal"`, or `"Ordinal"`. |
| `w`      | Initial interval width (scalar or list per block). |

```{r slice}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "Slice",
  Specs     = list(B = NULL,
                   Bounds = c(-Inf, Inf),
                   m = 100,
                   Type = "Continuous",
                   w = 1))
```

## Automated Factor Slice Sampler (`AFSS`)

**Description.**  Tibbits et al. (2014) proposed AFSS, which adapts an
orthogonal factor structure of the parameter space to align the slice
with the principal axes of the posterior.  This greatly reduces the
number of steps needed to create the slice in correlated,
high-dimensional problems.

**Specs.**

| Spec | Description |
|------|-------------|
| `A` | Adaptive burn-in iterations to discard. |
| `B` | Optional blocks. |
| `m` | Maximum steps for growing slice interval. |
| `n` | Previous iterations. |
| `w` | Initial interval width. |

```{r afss}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "AFSS",
  Specs     = list(A = Inf, B = NULL, m = 100, n = 0, w = 1))
```

## Univariate Eigenvector Slice Sampler (`UESS`)

**Description.**  Thompson (2011) proposed slicing along the eigenvectors
of the empirical covariance matrix of the chain.  Moving along
decorrelated directions yields a slice interval that is better aligned
with the posterior geometry.

**Specs.**

| Spec | Description |
|------|-------------|
| `A` | Adaptive burn-in iterations to discard. |
| `B` | Optional list of parameter blocks. |
| `m` | Maximum steps for the slice interval. |
| `n` | Number of previous iterations (for continuing a run). |

```{r uess}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "UESS",
  Specs     = list(A = Inf, B = NULL, m = 100, n = 0))
```

## Reflective Slice Sampler (`RSS`)

**Description.**  An extension of the slice sampler that reflects the
proposal off the boundary of the slice rather than shrinking, following
the trajectory of a billiard ball.  This can give better mixing in
elongated or curved posteriors.

**Specs.**

| Spec | Description |
|------|-------------|
| `m` | Steps per iteration. |
| `w` | Step size (scalar or parameter-length vector). |

```{r rss}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "RSS",
  Specs     = list(m = 1, w = 1))
```

## Oblique Hyperrectangle Slice Sampler (`OHSS`)

**Description.**  Generalises the hyperrectangle slice sampler to oblique
(rotated) hyperrectangles, allowing better coverage of posteriors with
axis-aligned correlations.

**Specs.**

| Spec | Description |
|------|-------------|
| `A` | Number of adaptive (warm-up) iterations; when exceeded the chain stops adapting. |
| `n` | Number of previous iterations (for continuing a run). |

```{r ohss}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "OHSS",
  Specs     = list(A = Inf, n = 0))
```

## Elliptical Slice Sampler (`ESS`)

**Description.**  Murray, Adams and MacKay (2010) proposed ESS for
models with a Gaussian prior.  A "ellipse" of proposals is defined
using the current state and a Gaussian draw; the algorithm searches the
ellipse for a point that satisfies the likelihood slice condition.  No
tuning is required.

**Specs.**

| Spec | Description |
|------|-------------|
| `B` | Optional blocks. |

```{r ess}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "ESS",
  Specs     = list(B = NULL))
```

---

# Miscellaneous and Specialised Algorithms

## t-walk (`twalk`)

**Description.**  Christen and Fox (2010) proposed the t-walk, a
self-contained, essentially tuning-free sampler that maintains two
points $\theta$ and $\phi$ and proposes moves using a random mixture of
four "kernels": walk, traverse, blow, and hop.  The traverse kernel
generates proposals along the line connecting the two points and is
particularly effective for correlated posteriors.

**Specs.**

| Spec | Description |
|------|-------------|
| `at` | Traverse move scale (recommended 6; range [2, 10]). |
| `aw` | Walk move scale (recommended 1.5; range [0.3, 2]). |
| `n1` | Subset size for traverse move (recommended 4; range [2, 20]). |
| `SIV` | Secondary initial values (must differ element-wise from `Initial.Values`). |

```{r twalk}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "twalk",
  Specs     = list(at = 6, aw = 1.5, n1 = 4, SIV = NULL))
```

## Refractive Sampler (`Refractive`)

**Description.**  Boyles and Welling (2012) proposed a sampler inspired
by Snell's law of refraction.  A particle crosses the boundary of a
"lens" defined by the current log-posterior value and refracts
according to the ratio of densities on either side.  This creates
proposals that preferentially explore high-density regions.

**Specs.**

| Spec | Description |
|------|-------------|
| `Adaptive` | Iteration at which adaptation begins. |
| `m`        | Steps per iteration (integer ≥ 2). |
| `w`        | Step size scalar. |
| `r`        | Refractive index ratio $r_1/r_2$. |

```{r refractive}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "Refractive",
  Specs     = list(Adaptive = 1, m = 2, w = 0.1, r = 1.3))
```

## Reversible-Jump (`RJ`)

**Description.**  Green's (1995) reversible-jump MCMC allows the
dimension of the parameter space to change between iterations, enabling
joint inference over model selection and parameters.  At each iteration
a parameter is either added to or removed from the active set, with a
binomial prior on model size.

**Specs.**

| Spec | Description |
|------|-------------|
| `bin.n`      | Binomial prior size $n$. |
| `bin.p`      | Binomial prior probability $p$. |
| `parm.p`     | Selection probabilities for each parameter (length $d$ vector). |
| `selectable` | Binary vector: 1 = can be selected, 0 = always included. |
| `selected`   | Binary vector indicating initial selection state. |

```{r rj}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "RJ",
  Specs     = list(bin.n = d, bin.p = 0.5,
                   parm.p   = rep(0.5, d),
                   selectable = rep(1, d),
                   selected   = rep(1, d)))
```

---

# Sequential / Updating Algorithms

These algorithms are designed for online updating as new data arrive,
using dynamic state-space models.  Non-dynamic parameters are updated
first, then dynamic parameters are updated sequentially through
time-periods.

## Sequential Metropolis-within-Gibbs (`SMWG`)

**Description.**  Applies MWG updates to dynamic state parameters in
a sequential state-space model.  The `Dyn` matrix specifies dynamic
parameter values for each of $T$ time-periods.

**Specs.**

| Spec | Description |
|------|-------------|
| `Dyn` | $T \times K_{\text{dyn}}$ matrix of dynamic parameter values. |

```{r smwg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "SMWG",
  Specs     = list(Dyn = Dyn_matrix))
```

## Sequential Adaptive Metropolis-within-Gibbs (`SAMWG`)

**Description.**  Extends SMWG with componentwise adaptive proposals for
dynamic parameters.

**Specs.**

| Spec | Description |
|------|-------------|
| `Dyn`         | Dynamic parameter matrix. |
| `Periodicity` | Adaptation frequency. |

```{r samwg}
Fit <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "SAMWG",
  Specs     = list(Dyn = Dyn_matrix, Periodicity = 50))
```

## Updating Sequential Metropolis-within-Gibbs (`USMWG`)

**Description.**  An *updating* variant of SMWG intended for online
inference.  Past samples stored in a `Fit` object are retained; only
the time-periods from `Begin` onwards are re-sampled.

**Specs.**

| Spec | Description |
|------|-------------|
| `Begin` | Time-period from which to begin updating. |
| `Dyn`   | Dynamic parameter matrix. |
| `Fit`   | Previous `demonoid` fit object. |

```{r usmwg}
Fit_update <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "USMWG",
  Specs     = list(Begin = T_new, Dyn = Dyn_matrix, Fit = Fit_old))
```

## Updating Sequential Adaptive Metropolis-within-Gibbs (`USAMWG`)

**Description.**  Combines the adaptive proposals of SAMWG with the
online updating capability of USMWG.

**Specs.**

| Spec | Description |
|------|-------------|
| `Begin`       | Start time-period for updating. |
| `Dyn`         | Dynamic parameter matrix. |
| `Fit`         | Previous `demonoid` fit. |
| `Periodicity` | Adaptation frequency. |

```{r usamwg}
Fit_update <- LaplacesDemon(Model, MyData, iv,
  Algorithm = "USAMWG",
  Specs     = list(Begin = T_new, Dyn = Dyn_matrix,
                   Fit = Fit_old, Periodicity = 50))
```

---

# Summary Table

The table below provides a quick reference for all available algorithms.
"Gradient?" indicates whether the log-posterior gradient is needed.
"Adaptive?" indicates built-in adaptation of the proposal.  "Parallel?"
indicates whether the algorithm natively uses multiple chains.

| Algorithm | Full name | Gradient? | Adaptive? | Parallel? |
|-----------|-----------|:---------:|:---------:|:---------:|
| `ADMG`    | Adaptive Directional MWG | No | Yes | No |
| `AGG`     | Adaptive Griddy-Gibbs | No | Yes | No |
| `AHMC`    | Adaptive HMC | Yes | Yes | No |
| `AIES`    | Affine-Invariant Ensemble Sampler | No | No | Yes |
| `AFSS`    | Automated Factor Slice Sampler | No | Yes | No |
| `AM`      | Adaptive Metropolis | No | Yes | No |
| `AMM`     | Adaptive-Mixture Metropolis | No | Yes | No |
| `AMWG`    | Adaptive MWG | No | Yes | No |
| `CHARM`   | Componentwise Hit-And-Run MH | No | Optional | No |
| `DEMC`    | Differential Evolution MC | No | No | Yes |
| `DRAM`    | Delayed Rejection AM | No | Yes | No |
| `DRM`     | Delayed Rejection MH | No | No | No |
| `ESS`     | Elliptical Slice Sampler | No | No | No |
| `GG`      | Griddy-Gibbs | No | No | No |
| `Gibbs`   | Gibbs Sampler | No | No | No |
| `HARM`    | Hit-And-Run MH | No | Optional | No |
| `HMC`     | Hamiltonian MC | Yes | No | No |
| `HMCDA`   | HMC with Dual-Averaging | Yes | Yes | No |
| `IM`      | Independence Metropolis | No | No | No |
| `INCA`    | Interchain Adaptation | No | Yes | Yes |
| `MALA`    | Metropolis-Adjusted Langevin | Yes | Yes | No |
| `MCMCMC`  | Metropolis-Coupled MC^3 | No | No | Yes |
| `MTM`     | Multiple-Try MH | No | No | No |
| `MWG`     | Metropolis-within-Gibbs | No | No | No |
| `NUTS`    | No-U-Turn Sampler | Yes | Yes | No |
| `OHSS`    | Oblique Hyperrectangle Slice | No | Yes | No |
| `pCN`     | Preconditioned Crank-Nicolson | No | No | No |
| `RAM`     | Robust Adaptive Metropolis | No | Yes | No |
| `RDMH`    | Random Dive MH | No | No | No |
| `Refractive` | Refractive Sampler | No | Yes | No |
| `RJ`      | Reversible-Jump | No | No | No |
| `RSS`     | Reflective Slice Sampler | No | No | No |
| `RWM`     | Random-Walk Metropolis | No | No | No |
| `SAMWG`   | Sequential Adaptive MWG | No | Yes | No |
| `SGLD`    | Stochastic Gradient Langevin | Yes | No | No |
| `Slice`   | Slice Sampler | No | No | No |
| `SMWG`    | Sequential MWG | No | No | No |
| `THMC`    | Tempered HMC | Yes | No | No |
| `twalk`   | t-walk | No | No | No |
| `UESS`    | Univariate Eigenvector Slice | No | Yes | No |
| `USAMWG`  | Updating Sequential Adaptive MWG | No | Yes | No |
| `USMWG`   | Updating Sequential MWG | No | No | No |

---

# Choosing an Algorithm

## No gradient available

- **Low dimension** ($d \lesssim 20$): `MWG`, `AMWG`, `CHARM`,
  `HARM` — simple, robust, easy to tune.
- **Moderate dimension** ($d \sim 20$–$100$): `AM`, `AMM`, `DRAM`,
  `RAM` — learn the proposal covariance adaptively.
- **Correlated parameters**: `AMM`, `DEMC`, `AIES`, `twalk` — use
  inter-chain information or direction-aware proposals.
- **Multimodal posterior**: `MCMCMC`, `THMC`, `DEMC` — tempering or
  parallel chains help escape local modes.
- **Discrete parameters**: `GG`, `AGG`, `Slice` — exact or grid-based
  sampling.
- **Variable selection**: `RJ` — reversible-jump for trans-dimensional
  moves.
- **Big data**: `SGLD` — stochastic gradients on minibatches.
- **Online updating / state-space models**: `SMWG`, `SAMWG`, `USMWG`,
  `USAMWG`.

## Gradient available

- **General purpose**: `NUTS` — nearly tuning-free, state-of-the-art
  gradient-based sampler.
- **Fixed hyperparameters**: `HMC` — highly efficient when $\epsilon$
  and $L$ are well tuned.
- **Big data**: `SGLD` — gradient estimated on minibatch.
- **Multimodal**: `THMC` — tempered momentum for better exploration.
- **Infinite-dimensional / functional**: `pCN` — dimension-robust.
- **Gaussian likelihood / prior**: `ESS` — requires no tuning.

## The default: `CHARM`

**Highlander** sets `CHARM` as its default `LaplacesDemon` algorithm
because it offers a good
balance of generality, robustness, and ease of use.  It requires no
gradient, adapts its scale automatically, and handles correlated
parameters better than standard `MWG` by mixing directions.  For most
problems it is a sensible starting point before switching to a more
specialised algorithm if needed.

---

# References

Atchadé, Y.F. (2006). "An Adaptive Version for the Metropolis Adjusted
Langevin Algorithm with a Truncated Drift". *Methodology and Computing in
Applied Probability*, 8, 235–254.

Bai, Y. (2009). "An Adaptive Directional Metropolis-within-Gibbs
Algorithm". Technical Report, University of Toronto.

Beskos, A., Roberts, G.O., Stuart, A.M., and Voss, J. (2008). "MCMC
Methods for Diffusion Bridges". *Stoch. Dyn.*, 8, 319–350.

Boyles, L.B. and Welling, M. (2012). "Refractive Sampling".

Christen, J.A. and Fox, C. (2010). "A General Purpose Sampling Algorithm
for Continuous Distributions (the t-walk)". *Bayesian Analysis*, 5(2),
263–282.

Craiu, R.V., Rosenthal, J., and Yang, C. (2009). "Learn From Thy
Neighbor: Parallel-Chain and Regional Adaptive MCMC". *Journal of the
American Statistical Association*, 104(488), 1454–1466.

Duane, S., Kennedy, A.D., Pendleton, B.J., and Roweth, D. (1987). "Hybrid
Monte Carlo". *Physics Letters B*, 195, 216–222.

Dutta, S. (2012). "Multiplicative Random Walk Metropolis-Hastings on the
Real Line". *Sankhya B*, 74(2), 315–342.

Gelman, A., Roberts, G.O., and Gilks, W.R. (1996). "Efficient Metropolis
Jumping Rules". *Bayesian Statistics*, 5, 599–607.

Geman, S. and Geman, D. (1984). "Stochastic Relaxation, Gibbs
Distributions, and the Bayesian Restoration of Images". *IEEE Transactions
on Pattern Analysis and Machine Intelligence*, 6(6), 721–741.

Geyer, C.J. (1991). "Markov Chain Monte Carlo Maximum Likelihood".
*Proceedings of the 23rd Symposium of the Interface*. Interface
Foundation. 156–163.

Goodman, J. and Weare, J. (2010). "Ensemble Samplers with Affine
Invariance". *Communications in Applied Mathematics and Computational
Science*, 5(1), 65–80.

Green, P.J. (1995). "Reversible Jump Markov Chain Monte Carlo Computation
and Bayesian Model Determination". *Biometrika*, 82, 711–732.

Haario, H., Laine, M., Mira, A., and Saksman, E. (2006). "DRAM: Efficient
Adaptive MCMC". *Statistical Computing*, 16, 339–354.

Haario, H., Saksman, E., and Tamminen, J. (2001). "An Adaptive Metropolis
Algorithm". *Bernoulli*, 7, 223–242.

Hastings, W.K. (1970). "Monte Carlo Sampling Methods Using Markov Chains
and Their Applications". *Biometrika*, 57(1), 97–109.

Hoffman, M.D. and Gelman, A. (2014). "The No-U-Turn Sampler: Adaptively
Setting Path Lengths in Hamiltonian Monte Carlo". *Journal of Machine
Learning Research*, 15, 1593–1623.

Liu, J., Liang, F., and Wong, W. (2000). "The Multiple-Try Method and
Local Optimization in Metropolis Sampling". *Journal of the American
Statistical Association*, 95, 121–134.

Metropolis, N., Rosenbluth, A.W., Rosenbluth, M.N., and Teller, E. (1953).
"Equation of State Calculations by Fast Computing Machines". *Journal of
Chemical Physics*, 21, 1087–1092.

Mira, A. (2001). "On Metropolis-Hastings Algorithms with Delayed
Rejection". *Metron*, LIX(3–4), 231–241.

Murray, I., Adams, R.P., and MacKay, D.J. (2010). "Elliptical Slice
Sampling". *Journal of Machine Learning Research*, 9, 541–548.

Neal, R.M. (2003). "Slice Sampling" (with discussion). *Annals of
Statistics*, 31(3), 705–767.

Neal, R.M. (2011). "MCMC Using Hamiltonian Dynamics". In Brooks, S.,
Gelman, A., Jones, G.L., and Meng, X.-L. (eds.), *Handbook of Markov
Chain Monte Carlo*. Chapman & Hall / CRC Press. Chapter 5.

Ritter, C. and Tanner, M. (1992). "Facilitating the Gibbs Sampler: the
Gibbs Stopper and the Griddy-Gibbs Sampler". *Journal of the American
Statistical Association*, 87, 861–868.

Roberts, G.O. and Rosenthal, J.S. (2009). "Examples of Adaptive MCMC".
*Computational Statistics and Data Analysis*, 18, 349–367.

Roberts, G.O. and Tweedie, R.L. (1996). "Exponential Convergence of
Langevin Distributions and Their Discrete Approximations". *Bernoulli*,
2(4), 341–363.

Smith, R.L. (1984). "Efficient Monte Carlo Procedures for Generating
Points Uniformly Distributed Over Bounded Region". *Operations Research*,
32, 1296–1308.

Ter Braak, C.J.F. and Vrugt, J.A. (2008). "Differential Evolution Markov
Chain with Snooker Updater and Fewer Chains". *Statistics and Computing*,
18(4), 435–446.

Thompson, M.D. (2011). "Slice Sampling with Multivariate Steps".
<https://utoronto.scholaris.ca/items/161206c5-af25-41a8-99d8-8d622f950cb3>

Tibbits, M., Groendyke, C., Haran, M., and Liechty, J. (2014).
"Automated Factor Slice Sampling". *Journal of Computational and Graphical
Statistics*, 23(2), 543–563.

Vihola, M. (2011). "Robust Adaptive Metropolis Algorithm with Coerced
Acceptance Rate". *Statistics and Computing*. Springer, Netherlands.

Welling, M. and Teh, Y.W. (2011). "Bayesian Learning via Stochastic
Gradient Langevin Dynamics". *Proceedings of the 28th International
Conference on Machine Learning (ICML)*, 681–688.
